---
  title: "Final Project"
output: html_document
---
  #https://www.kaggle.com/carolzhangdc/predict-imdb-score-with-data-mining-algorithms
  
  ```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
library(forecast)
library(leaps)
library(forecast)
library(caret)
library(car)
library(data.table)
library(VIM)
library(corrplot)
library(ggplot2)
library(ggcorrplot)
library(plotly)
library(ggrepel)
library(caret)
library(ggthemes)
library(rpart)				        # Popular decision tree algorithm
library(rattle)					# Fancy tree plot
```


# Read Datasets and removing duplicate datas
```{r, warning=FALSE, message=FALSE}
library(readr)
movie_metadata <- read_csv("movie_metadata.csv")
sum(duplicated(movie_metadata))
movie_metadata <- movie_metadata[!duplicated(movie_metadata),] #removing duplicate row
str(movie_metadata)
#dealing with missing data 
sum(is.na(movie_metadata)) #2674 of null values
colSums(sapply(movie_metadata, is.na))
missing.values <- aggr(movie_metadata, sortVars = T, prop = T, sortCombs = T, cex.lab = 1.5, cex.axis = .6, cex.numbers = 5, combined = F, gap = -.2) #visualize missing value
#Gross is missing 17% or data and budget is 9% of data, hence we just have to remove them
movie_metadata <- movie_metadata[!is.na(movie_metadata$gross),]
movie_metadata <- movie_metadata[!is.na(movie_metadata$budget),]
colSums(sapply(movie_metadata, is.na))
missing.values <- aggr(movie_metadata, sortVars = T, prop = T, sortCombs = T, cex.lab = 1.5, cex.axis = .6, cex.numbers = 5, combined = F, gap = -.2) #visualize missing value
# aspect ratio still has 74 missing value, lets inspect that
table(movie_metadata$aspect_ratio)
movie_metadata$aspect_ratio[is.na(movie_metadata$aspect_ratio)] <- 0 # creplacing the null aspect ratio with 0 
mean(movie_metadata$gross[movie_metadata$aspect_ratio == 1.85]) #checking the mean of gross 
mean(movie_metadata$gross[movie_metadata$aspect_ratio == 2.35])
mean(movie_metadata$gross[movie_metadata$aspect_ratio != 1.85 & movie_metadata$aspect_ratio != 2.35])
mean(movie_metadata$gross[movie_metadata$aspect_ratio == 0]) # We can tell that with aspect ratio being null value the gross will be significantly less. We will keep this as is aspect ratio has a hiarchical structure. The higher Aspecct ratio the higher the gross are. 
summary(movie_metadata$gross)
# deal with other missing variable
colSums(sapply(movie_metadata, is.na)) #lets start with content rating 
table(movie_metadata$content_rating)
movie_metadata <-movie_metadata [!is.na(movie_metadata$content_rating),] # removing null value in movie_metadata
sum(is.na(movie_metadata$content_rating)) # double checking the sum of missing value
table(movie_metadata$content_rating)
#replacing rating with the modern rating
movie_metadata$content_rating[movie_metadata$content_rating == "M"] <- "PG"
movie_metadata$content_rating[movie_metadata$content_rating == "GP"] <- "PG"
movie_metadata$content_rating[movie_metadata$content_rating == "X"] <- "NC-17"
table(movie_metadata$content_rating)
#replacing the rest of rating with the most common rating of R
movie_metadata$content_rating[movie_metadata$content_rating == "Approved"] <- "R"
movie_metadata$content_rating[movie_metadata$content_rating == "Not Rated"] <- "R"
movie_metadata$content_rating[movie_metadata$content_rating == "Passed"] <- "R"
movie_metadata$content_rating[movie_metadata$content_rating == "Unrated"] <- "R"
table(movie_metadata$content_rating)
# Now lets look the rest
colSums(sapply(movie_metadata, is.na)) 
# replace NA with column average for facenumber_in_poster
movie_metadata$facenumber_in_poster[is.na(movie_metadata$facenumber_in_poster)] <- round(mean(movie_metadata$facenumber_in_poster, na.rm = TRUE))
# convert 0s into NAs for other predictors
movie_metadata[,c(5,6,8,13,24,26)][movie_metadata[,c(5,6,8,13,24,26)] == 0] <- NA
# impute missing value with column mean
movie_metadata$num_critic_for_reviews[is.na(movie_metadata$num_critic_for_reviews)] <- round(mean(movie_metadata$num_critic_for_reviews, na.rm = TRUE))
movie_metadata$duration[is.na(movie_metadata$duration)] <- round(mean(movie_metadata$duration, na.rm = TRUE))
movie_metadata$director_facebook_likes[is.na(movie_metadata$director_facebook_likes)] <- round(mean(movie_metadata$director_facebook_likes, na.rm = TRUE))
movie_metadata$actor_3_facebook_likes[is.na(movie_metadata$actor_3_facebook_likes)] <- round(mean(movie_metadata$actor_3_facebook_likes, na.rm = TRUE))
movie_metadata$actor_1_facebook_likes[is.na(movie_metadata$actor_1_facebook_likes)] <- round(mean(movie_metadata$actor_1_facebook_likes, na.rm = TRUE))
movie_metadata$cast_total_facebook_likes[is.na(movie_metadata$cast_total_facebook_likes)] <- round(mean(movie_metadata$cast_total_facebook_likes, na.rm = TRUE))
movie_metadata$actor_2_facebook_likes[is.na(movie_metadata$actor_2_facebook_likes)] <- round(mean(movie_metadata$actor_2_facebook_likes, na.rm = TRUE))
movie_metadata$movie_facebook_likes[is.na(movie_metadata$movie_facebook_likes)] <- round(mean(movie_metadata$movie_facebook_likes, na.rm = TRUE))
colSums(sapply(movie_metadata, is.na)) #check out the result (We still have some left)
# See does language matter 
table(movie_metadata$language) # We can tell mostly is English, hence it shouldn't matter to us.
movie_metadata <- movie_metadata[,-c(20) ] # We drop language column as a result
# See does Color matter
colSums(sapply(movie_metadata, is.na)) 
table(movie_metadata$color)
movie_metadata <- movie_metadata[,-1] # It does not matter, hence we decided to drop it. 
# There are 30 more missing value (Since it is so small (< 1%) compare to our sample, we decided to drop it.)
colSums(sapply(movie_metadata, is.na)) 
sum(is.na(movie_metadata))
movie_metadata <- na.omit(movie_metadata)
colSums(sapply(movie_metadata, is.na)) 
# Now lets look at language 
table(movie_metadata$country) # We can tell most movie are from US and UK, we will create three catergory. (US, UK and Other)
movie_metadata$country[movie_metadata$country != "USA"  & movie_metadata$country !="UK"] <- "Others"
table(movie_metadata$country) 
```

# Tidy up movie Title
```{r}
library(stringr)
movie_metadata$movie_title <- gsub("Â", "", as.character(factor(movie_metadata$movie_title)))
str_trim(movie_metadata$movie_title, side = "right")
```

# Work on Genres
```{r}
# create a new data frame
genres.df <- as.data.frame(movie_metadata[,c("genres", "gross")])
# separate different genres into new columns
genres.df$Action <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Action") 1 else 0)
genres.df$Adventure <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Adventure") 1 else 0)
genres.df$Animation <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Animation") 1 else 0)
genres.df$Biography <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Biography") 1 else 0)
genres.df$Comedy <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Comedy") 1 else 0)
genres.df$Crime <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Crime") 1 else 0)
genres.df$Documentary <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Documentary") 1 else 0)
genres.df$Drama <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Drama") 1 else 0)
genres.df$Family <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Family") 1 else 0)
genres.df$Fantasy <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Fantasy") 1 else 0)
genres.df$`Film-Noir` <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Film-Noir") 1 else 0)
genres.df$History <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "History") 1 else 0)
genres.df$Horror <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Horror") 1 else 0)
genres.df$Musical <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Musical") 1 else 0)
genres.df$Mystery <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Mystery") 1 else 0)
genres.df$News <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "News") 1 else 0)
genres.df$Romance <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Romance") 1 else 0)
genres.df$`Sci-Fi` <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Sci-Fi") 1 else 0)
genres.df$Short <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Short") 1 else 0)
genres.df$Sport <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Sport") 1 else 0)
genres.df$Thriller <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Thriller") 1 else 0)
genres.df$War <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "War") 1 else 0)
genres.df$Western <- sapply(1:length(genres.df$genres), function(x) if (genres.df[x,1] %like% "Western") 1 else 0)
# get the mean of Gross for different genres
means <- rep(0,23)
for (i in 1:23) {
  means[i] <- mean(genres.df$gross[genres.df[i+2]==1])
}
head(means)
# plot the means
barplot(means/10^6, main = "Average gross for different genres")
head(genres.df$action)
# We found out genre does matter to the Gross, hence we combine both df
movie_metadata <- cbind(movie_metadata, genres.df)
movie_metadata <- movie_metadata[, -c(9, 27, 28)]
```
# looking at movie release year histogram
```{r}
hist(movie_metadata$title_year) # find out movie release before 1980 is probably irrelevent, hence we remove any movie that is release before 1980
movie_metadata <- movie_metadata[movie_metadata$title_year>= 1980, ] 
```
# Little visualization before predicting
```{r}
movie_metadata %>%
  plot_ly(x = ~movie_facebook_likes, y = ~gross, color = ~content_rating , mode = "markers", text = ~content_rating, alpha = 0.7, type = "scatter")
movie_metadata$ROI <- round(movie_metadata$gross / movie_metadata$budget *100, 2)
movie_metadata$profitable <- ifelse(movie_metadata$ROI > 1, 1, 0)
summary(movie_metadata$ROI)
summary(movie_metadata$profit)
summary(movie_metadata$gross)
```


```{r}
# Looking at the top 25 gross movie ROI
movie_metadata %>% 
  filter(budget > 100000) %>% 
  arrange(desc(gross)) %>%
  top_n(25, gross) %>%
  ggplot(aes(x = budget/10^6, y= ROI)) + geom_point() + geom_smooth() + geom_text_repel(aes(label = movie_title), size = 3) +
  labs(x = "Budget in Millions ($)", y = "ROI (%)", title = "Top 25 movie ROI base on gross" )+ 
  theme_economist() + 
  scale_color_economist()
```


# Checking out the does the name of director and actors matter
```{r}
uniqueN(movie_metadata$director_name)
uniqueN(movie_metadata$actor_1_name)
uniqueN(movie_metadata$actor_3_name)
uniqueN(movie_metadata$actor_2_name)
uniqueN(movie_metadata$plot_keywords)
# all of them are all different, hence it make no sense to use them to predict. We decided to drop movie title, plot_keywords and movie_imdb_link too. 
final_df <- subset(movie_metadata, select = -c(director_name, actor_2_name, actor_1_name, movie_title, actor_3_name, plot_keywords, movie_imdb_link))
```

# Change catergorical variable to catergorical and finalize the final_df
```{r, fig.height= 10}
#Checking for genre that is not usable and drop it
sum(uniqueN(final_df$Action))
sum(uniqueN(final_df$Adventure))
sum(uniqueN(final_df$Animation))
sum(uniqueN(final_df$Biography))
sum(uniqueN(final_df$Comedy))
sum(uniqueN(final_df$Crime))
sum(uniqueN(final_df$Documentary))
sum(uniqueN(final_df$Drama))
sum(uniqueN(final_df$Family))
sum(uniqueN(final_df$Fantasy))
sum(uniqueN(final_df$`Film-Noir`)) # Remove
sum(uniqueN(final_df$History))
sum(uniqueN(final_df$Horror))
sum(uniqueN(final_df$Musical))
sum(uniqueN(final_df$Mystery))
sum(uniqueN(final_df$News)) # Remove
sum(uniqueN(final_df$Romance))
sum(uniqueN(final_df$`Sci-Fi`))
sum(uniqueN(final_df$Short)) # Remove
sum(uniqueN(final_df$Thriller))
sum(uniqueN(final_df$War))
sum(uniqueN(final_df$Western))
final_df <- subset(final_df, select = -c(`Film-Noir`, News, Short))
# factor country and content rating
final_df$country <- (as.factor(final_df$country))
final_df$content_rating <- (as.factor(final_df$content_rating))
str(final_df)
final_df <- subset(final_df, select = -c(ROI, profitable)) #removing ROI and Profitable column
# try basic lm
reg1 <- lm(gross ~ . , data = final_df)
summary(reg1)
```
# Checking for final_df correlation
```{r, fig.height=10, fig.width=10}
final_df_matrix <- as.matrix(final_df[, c(-11,-12,-39)]) #getting rid of factor variable (country, content rating and gross_category)
m <- cor(final_df_matrix)
corrplot(m, method = "circle", type = "upper", order = "hclust")
```

# Start predicting 
```{r}
# Cut gross in to percentile (cut it by 0 to 25 percentile,25 to 75 percentile, Above 75 percentile)
summary(final_df$gross)
final_df$gross_catogorical <- cut(final_df$gross/10^6, breaks = c(0, 7, 66, 761)) # in millions
sum(is.na(final_df$gross_catogorical)) # double checking the completeness of data
# partition data
set.seed(1)  # set seed for reproducing the partition
train.index <- sample(c(1:3686), 3686*0.6)  
#Create and set aside the remaining 40% of the data, to be used after omitting unhelpful data points and unnecessary variables.
train.df <- final_df[train.index,]
valid.df <- final_df[-train.index,]
library(rpart)
```

# Linear model (Don't use this model, since our target variable are now factor variable)
```{r, fig.height= 10}
reg2 <- lm(gross ~ . , data = train.df)
summary(reg2)
vif(reg2)
reg2 <- lm(gross ~ . -actor_1_facebook_likes -cast_total_facebook_likes -actor_2_facebook_likes -gross_catogorical , data = train.df) # Removed colinear variables
reg.step <- step(reg2, direction = "both")
summary(reg.step)
vif(reg.step)
reg.step.pred <- predict(reg.step, valid.df)
accuracy(reg.step.pred, valid.df$gross)
par(mfrow=c(2,2))
plot(reg.step)
all.residuals <- (valid.df$gross - reg.step.pred)/10^6
hist(all.residuals, breaks = 25, xlab = "Residuals", main = "")
data.frame("Predicted" = reg.step.pred, "Actual" = valid.df$gross,
           "Residual" = all.residuals)[0:20,]
```
Final Linear model with a R-square of 0.5734 




# Classification tree
```{r}
library(rpart)
library(rpart.plot)
class.tree <- rpart(gross_catogorical ~. -gross, data = train.df, method = "class")
prp(class.tree, type = 1, extra = "auto", split.font = 1, varlen = 0)
fancyRpartPlot(class.tree)
class.tree.pred <- predict(class.tree, valid.df, type = "class")
#accuracy(class.tree, valid.df$gross)
confusionMatrix(class.tree.pred, as.factor(valid.df$gross_catogorical))
```
Classification tree accuracy = 0.6922

# Pruned Tree
```{r}
# Cross Validation
set.seed(2)
crossvalid_ct <- rpart(gross_catogorical ~. - gross, data = train.df, method = "class", cp = 0.001, minsplit = 5, xval = 5)
printcp(crossvalid_ct)
```





```{r}
# prune tree with the smallest tree within 1 xstd of min. error (13 split is the best)
pruned_ct <- prune(crossvalid_ct, 
                   cp = 0.0069897)
length(pruned_ct$frame$var[pruned_ct$frame$var == "<leaf>"])
prp(pruned_ct)
fancyRpartPlot(pruned_ct)
```
# Apply prune tree model with prediction
```{r}
pruned_ct_pred <- predict(pruned_ct, train.df, type = "class")
confusionMatrix(pruned_ct_pred, train.df$gross_catogorical)
pruned_ct_pred <- predict(pruned_ct, valid.df, type = "class")
confusionMatrix(pruned_ct_pred, valid.df$gross_catogorical)
```
Accuracy for testing dataset is 0.7752
Accuracy for validation dataset is 0.6949

## K-Nearest Neighbors
# normalize our data
```{r}
library(FNN)
# initialize normalized training, validation, test data, complete data frames to originals
train.norm <- train.df
valid.norm <- valid.df
# use preProcess() from the caret package to normalize predictors.
norm.values <- preProcess(train.df[, -39-6], method=c("center", "scale"))
train.norm[, -39-6] <- predict(norm.values, train.df[, -39-6])
valid.norm[, -39-6] <- predict(norm.values, valid.df[, -39-6])
```



# remove column 'country', 'country rating', 'gross' (becuase KNN can only process numeric data, while 'country' and 'country rating' are factor variables.)
```{r}
valid.norm <- valid.norm[ -c(11,12,6) ]
train.norm <- train.norm[ -c(11,12,6) ]
```



#Find the best k

```{r}
# initialize a data frame with two columns: k, and accuracy.
accuracy.df <- data.frame(k = seq(1, 36, 1), accuracy = rep(0, 36))
# compute knn for different k on validation data.
for(i in 1:36) {
  knn.pred <- knn(train.norm[, -36], valid.norm[, -36],
                  cl = train.norm[, 36], k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.pred, valid.norm[, 36])$overall[1]
}
accuracy.df
```



#plot accuracy for each K
```{r}
ggplot(accuracy.df, aes(y = accuracy, x = k)) + geom_point() + geom_line()
knn.pred <- knn(train.norm[, -36], valid.norm[, -36],
                cl = train.norm[, 36], k = 14)
confusionMatrix(knn.pred, valid.norm[, 36])
#From the accuracy result, we found that the best k is 5. However, the highest accuracy is only 0.438
```
From the accuracy result, we found that the best k is 14. However, the highest accuracy is only 0.6122

```{r}
library(ggthemes)
# Change point shapes and line types by groups
ggplot(accuracy.df, aes(y = accuracy, x = k))+ 
  geom_point(color="yellow")+
  labs(title="Plot Accuracy for each K Value",x="K Value", y = "Accuracy")+
  theme_economist() + 
  scale_color_economist()+
  geom_step(color="yellow")
```


# random Forest

```{r}
library(randomForest)
set.seed(5)
# I removed 'Sci-Fi' column because the system said it could not find this column.
rf <- randomForest(gross_catogorical ~ . -gross -`Sci-Fi`- Documentary - Western - Musical - History - Sport - War - Biography - Mystery - Animation - Fantasy, data = train.df, mtry = 10, ntree = 1500)
# Show model error (We can see 500 tree is sufficient enough)
plot(rf)
legend('topright', colnames(rf$err.rate), col=1:4, fill=1:4)
#tuning mtry ( Number of variables randomly sampled as candidates at each split)
tune_rf <- tuneRF(train.df[,-c(39,6)], train.df[,39], mtrystart = 5, ntreeTry = 500, stepFactor=1.5, trace = TRUE, plot = TRUE, dobest = TRUE)
```

#install packages for further steps
```{r}
library(ggthemes)
```
#relative variable importance by plotting the mean decrease in Gini calculated across all trees
```{r}
# Get importance
importance <- importance(rf)
# ??? what is MeanDcreaseGini???
varImportance <- data.frame(Variables = row.names(importance), 
                            Importance = round(importance[ ,'MeanDecreaseGini'],2))
# Create a rank variable based on importance
rankImportance <- varImportance %>%
  mutate(Rank = paste0('#',dense_rank(desc(Importance))))
# Use ggplot2 to visualize the relative importance of variables
ggplot(rankImportance, aes(x = reorder(Variables, Importance), 
                           y = Importance, fill = Importance)) +
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
            hjust=0, vjust=0.55, size = 4, colour = 'red') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_few()
```
# Apply Model
```{r}
set.seed(632)
# apply model on validation set
rf.pred.valid <- predict(rf, valid.df)
# generate confusion matrix for validation data
confusionMatrix(rf.pred.valid, valid.df$gross_catogorical)
```
Random Forest accuracy = 0.7512

# Try predict profitability with GLM model
```{r}
# Create data frame with profit
final_df_profit <- final_df
final_df_profit$profitable <- as.factor(ifelse(final_df$gross-final_df$budget > 0, 1 ,0))
# partition data
set.seed(3)  # set seed for reproducing the partition
train.index <- sample(c(1:3686), 3686*0.6)  
#Create and set aside the remaining 40% of the data, to be used after omitting unhelpful data points and unnecessary variables.
train.df.logistic <- final_df_profit[train.index,]
valid.df.logistic <- final_df_profit[-train.index,]
# Try using logistic model to predict wether a movie will profit or not (Profit = 1, Did not profit = 0)
logistic_reg <- glm(profitable~. - gross - gross_catogorical, data = train.df.logistic, family = "binomial")
summary(logistic_reg)
vif(logistic_reg)
logistic_reg <- glm(profitable ~. - gross - gross_catogorical - facenumber_in_poster - content_rating - aspect_ratio -movie_facebook_likes - Action -Adventure -Animation -Biography -History - Mystery - Sport - Thriller - War - Western - actor_1_facebook_likes - actor_2_facebook_likes - cast_total_facebook_likes - actor_3_facebook_likes - num_user_for_reviews - duration - Musical, data = train.df.logistic, family = "binomial")
summary(logistic_reg)
vif(logistic_reg)
#Apply logistics regression model
logistic_reg_pred <- predict(logistic_reg, valid.df.logistic[, - 40], type = "response")
confusionMatrix(as.factor(ifelse(logistic_reg_pred > 0.5 ,1, 0)), as.factor(valid.df.logistic$profitable))
```
Logistic regression Accuracy = 0.718


```{r}
logistic_reg <- glm(profitable ~. - gross - gross_catogorical - facenumber_in_poster - content_rating - aspect_ratio -movie_facebook_likes - Action -Adventure -Animation -Biography -History - Mystery - Sport - Thriller - War - Western - actor_1_facebook_likes - actor_2_facebook_likes - cast_total_facebook_likes - actor_3_facebook_likes - num_user_for_reviews - duration - Musical, data = train.df.logistic[-c(647,2817),], family = "binomial")
summary(logistic_reg)
vif(logistic_reg)
```


#Apply logistics regression model
```{r}
logistic_reg_pred <- predict(logistic_reg, valid.df.logistic[, - 40], type = "response")
confusionMatrix(as.factor(ifelse(logistic_reg_pred > 0.5 ,1, 0)), as.factor(valid.df.logistic$profitable))
```

